---
title: "Working with nflfastR rosters"
author: "Ben Baldwin"
date: "9/24/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

At long last, there's a way to merge the new play-by-play data with roster information in nflfastR 3.0. First, install the upgrade from CRAN:

``` {r eval = FALSE}
install.packages("nflfastR")
```

And then load the library:

``` {r}
library(nflfastR)
library(tidyverse)
```

The easy part is getting the rosters. There's a new function called `fast_scraper_roster`.

``` {r roster}
roster <- fast_scraper_roster(2019)

roster %>%
  filter(team == "SEA", position %in% c("WR", "TE"))
```

Now let's load play-by-play data from 2019:
``` {r roster_pbp_load}
games_2019 <- readRDS(url('https://raw.githubusercontent.com/guga31bb/nflfastR-data/master/data/play_by_play_2019.rds'))
```

Here is what the new player IDs look like:

``` {r roster_pbp}
games_2019 %>%
  filter(rush == 1 | pass == 1, posteam == "SEA") %>%
  select(desc, name, id)
```

But these IDs aren't very useful. So we need to decode them using the new function `decode_player_ids`:

``` {r roster_decode_ex}
games_2019 %>%
  filter(rush == 1 | pass == 1, posteam == "SEA") %>%
  nflfastR::decode_player_ids() %>%
  select(desc, name, id)
```

So now we have the familiar GSIS IDs. Let's apply this to the whole dataframe:

``` {r pbp_decode}
decoded_pbp <- games_2019 %>%
  nflfastR::decode_player_ids()
```

Now we're ready to join to the roster data using these IDs:
``` {r decode_join}
joined <- decoded_pbp %>% 
  filter(!is.na(receiver_id)) %>%
  select(posteam, season, desc, receiver, receiver_id, epa) %>%
  left_join(roster, by = c('receiver_id' = 'gsis_id'))
```

Now we can do something like look at the most receiving EPA by the top 5 players at each position group:
``` {r decode_table}
#the real work is done, this just makes a table and has it look nice
joined %>%
  filter(position %in% c('WR', 'TE', 'RB')) %>%
  group_by(receiver_id, receiver, position) %>%
  summarize(tot_epa = sum(epa), n=n()) %>%
  arrange(-tot_epa) %>%
  ungroup() %>%
  group_by(position) %>%
  mutate(position_rank = 1:n()) %>%
  filter(position_rank <= 5) %>%
  dplyr::rename(Pos_Rank = position_rank, Player = receiver, Pos = position, Tgt = n, EPA = tot_epa) %>%
  select(Player, Pos, Pos_Rank, Tgt, EPA) %>%
  knitr::kable(digits = 0)
```

Not surprisingly, all 5 of the top 5 WRs in terms of EPA added come in ahead of the top RB. Note that the number of targets won't match official stats because we're including plays with penalties.
